---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---

# Packages
Load packages, functions, & data; format data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(wesanderson)
library(lubridate)
library(readxl)
source("scripts/fishsum_function.R")
`%notin%` <- Negate(`%in%`)

# read in tagging data
# note: currently excludes herring w/ "226000011055/35/85" tag

tags <- read_xlsx(
  "other/1 IFW Tagging Data Golden Master 2_25_22.xlsx",
  sheet = "All Tagged Fish") %>%
  rename(tag = 'Tag Number',
         datetime = as.character('Date and Time'),
         Common.Name = 'Common Name') %>%
  mutate(tag = as.numeric(tag)) %>%
  filter(is.na(tag) == FALSE) %>%
  mutate(tag_short = str_sub(tag, - 9, - 1),
         tag_time = ymd_hms(datetime, 
                           tz = "America/New_York"))

# video time series
v_time <- 
  read_xlsx("other/2 IFW Video Data Golden Master 11_2_21.xlsx",
               sheet = "Grouped") %>%
  select(Date, time = Time, species = Species) %>%
  mutate(year = year(Date),
         # note: time read in as decimal hours (a prop. of 24 hr)
         time = as.numeric(as.character(time))*24, 
         hour = trunc(time),
         daytime = case_when(time/24 > (7/24) & 
                               time/24 < (19/24) ~ 1,
                             TRUE ~ 0)) %>%
  group_by(species, year, daytime) %>%
  tally() %>%
  filter(species %in% c("American Shad", "River Herring spp.")) %>%
  arrange(species, year, daytime) %>%
  group_by(species, year) %>%
  mutate(pday = n/sum(n))
# write.csv(v_time, here("output", 
#                           "proportion_counts_in_daytime.csv"), 
#           row.names=F)

v_temp <- v_time %>%
  group_by(species, year) %>%
  summarise(n = sum(n))

# historic Island Farm Wier Video counts
shad_hist <- data.frame(species = "American Shad",
                        year = c(1996:2005),
                        n = c(48, 76, 149, 582, NA, 717, 
                                1114, 362, NA, 13))

herring_hist <- data.frame(species = "River Herring spp.",
                           year = c(1996:2005),
                        n = c(0, 117, 9, 114, NA, 61, 
                                17, 12, NA, 0))

# combine recent and historic video count data
v <- v_temp %>%
  bind_rows(shad_hist, herring_hist) %>%
  arrange(species, year)

rm(v_temp)

# Compile receiver data
# Data are compiled, filtered, and cleaned within these scripts. Anthony should review to ensure all filtering was OK.
source("scripts/compile_pings_2012.R")
source("scripts/compile_pings_2013.R")
source("scripts/compile_pings_2014.R")
source("scripts/compile_pings_2015.R")
source("scripts/compile_pings_2016.R")
source("scripts/compile_pings_2017.R")
source("scripts/compile_pings_2018.R")
source("scripts/compile_pings_2019.R")

# Extracting & summarizing fish tag data
fish2012_sum <- fishsum(data2012)
fish2013_sum <- fishsum(data2013)
fish2014_sum <- fishsum(data2014)
fish2015_sum <- fishsum(data2015)
fish2016_sum <- fishsum(data2016)
fish2017_sum <- fishsum(data2017)
fish2018_sum <- fishsum(data2018)
fish2019_sum <- fishsum(data2019)
fish_sum <- fish2019_sum %>%
  bind_rows(fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum) %>%
  arrange(species, tag_time)

# remove unneeded objects
rm(fish2019_sum, fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum)

# subset target species and create 1/0 passage variable (did the fish reach A4)
f <- tags %>%
  select(tag_short, Common.Name, tag_time) %>%
  filter(Common.Name %in% c("American Shad",
                            "River Herring",
                            "Blueback Herring",
                            "Alewife")) %>%
  left_join(select(fish_sum, year = detect_year, tag_short, passage.days), by = "tag_short") %>%
  # remove 1 herring tagged in 2014 but only detected in 2015
  filter(is.na(passage.days) | passage.days < 300) %>%
  mutate(passage = case_when(is.na(passage.days) ~ 0,
                               TRUE ~ 1),
         year = as.factor(case_when(is.na(year) ~ year(tag_time),
                          TRUE ~ year)),
         year_num = as.numeric(as.factor(year)),
         year_char = paste0("y", year),
         sp = as.factor(case_when(grepl(Common.Name, 
                                        pattern="Herring|Alewife")~
                          "Herring",
                        TRUE ~ "Shad")))

shad_df <- f %>%
  filter(sp == "Shad")

herring_df <- f %>%
  filter(sp == "Herring")

# examine shad data
shad_raw_sum <- cbind.data.frame(
  trials = aggregate(shad_df$passage, 
                     by = list(shad_df$year), length),
  successes = aggregate(shad_df$passage, 
                        by = list(shad_df$year), sum)[,2],
  prop = aggregate(shad_df$passage, by = list(shad_df$year), mean)[,2]
  ) %>%
  mutate(year = as.numeric(as.character(trials.Group.1)),
         species = "American Shad") %>%
  select(-trials.Group.1)

# examine herring data
herring_raw_sum <- cbind.data.frame(
  trials = aggregate(herring_df$passage, 
                     by = list(herring_df$year), length),
  successes = aggregate(herring_df$passage, 
                        by = list(herring_df$year), sum)[,2],
  prop = aggregate(herring_df$passage, by = list(herring_df$year), mean)[,2]
  ) %>%
  mutate(year = as.numeric(as.character(trials.Group.1)),
         species = "Herring spp.") %>%
  select(-trials.Group.1)

raw_sum <- rbind(shad_raw_sum, herring_raw_sum)

shad_data <- list(
  D = 1,
  N = nrow(filter(f, Common.Name == "American Shad")),
  L = length(unique(filter(f, Common.Name == "American Shad")$year)),
  y = filter(f, Common.Name == "American Shad")$passage,
  ll = as.numeric(as.factor(filter(f, Common.Name == "American Shad")$year)),
  x = rbind(rep(1, nrow(filter(f, Common.Name == "American Shad"))))
)

# Format fish data for "passage proportion" analysis

# combine raw data summaries
both <- rbind(shad_raw_sum, herring_raw_sum) %>%
  select(N = trials.x, z = successes, -prop, year) %>%
  mutate(year = as.numeric(as.factor(year)),
         fish = c(rep(1,7), rep(2,8)),
         fishyears = 1:15) 

nfishyears <- 15
nyears <- 8
nfish <- 2

# bundle data for jags
fish_jags_data <- list(
  nfish = nfish,
  nfishyears = nfishyears,
  z = both$z,
  N = both$N,
  c = both$fish,
  fallback_detect = c(rep(1,33), rep(0,106)),
  fallback_n = 139
)
```
# Save data and diagnostic plot files for inspection
Need to be checked by Anthony.
```{r}
# source("scripts/save_diagnostics.R")
```
# JAGS model for fish passage proportion
logit-normal with hyperparameter for each fish
```{r}

cat(file = "scripts/fish_mod_logit.txt",
"
model{
for(s in 1:nfishyears){
  z[s] ~ dbin(ilogit(lp[s]), N[s])
  lp[s] ~ dnorm(mu_logit_p[c[s]], tau_logit_p[c[s]] )
  logit(p[s]) <- lp[s]
}  

for( c in 1:nfish){
mu_logit_p[c] ~ dnorm(0, 0.001)
tau_logit_p[c] ~ dgamma(0.001, 0.001)
logit(mu_p[c]) <- mu_logit_p[c]
}

fallback_p ~ dunif(0,1) # dbeta(3,3)
for(i in 1:length(fallback_detect)){
fallback_detect[i] ~ dbern(fallback_p)
}

# corrected for imperfect detection in the arrays
fallback_p_cor <- fallback_p / 0.5


}
"
)
```
# Run model
```{r}
fish_modn <- jagsUI::jags(
  data = fish_jags_data,
  model.file = "scripts/fish_mod_logit.txt",
  n.iter = 600000,
  n.burnin = 400000,
  parameters.to.save = c(
    "theta",
    "omega",
    "kappa",
    "S0",
    "R0",
    "p",
    "mu_p",
    "mu_logit_p",
    "tau_logit_p",
    "lp",
    "fallback_p",
    "fallback_p_cor"
  ),
  n.chains = 3,
  parallel = T,
  n.thin = 200,
  inits = function()
    list(
      lp = c(rnorm(7, 0, 0.1),
             rnorm(8, -10, 0.1)),
      omega =
        c(rnorm(1, .47, .01),
          rnorm(1, .02, .01)),
      kappaMinusTwo = c(rnorm(1, 400, 2),                                                      rnorm(1, 1000, 2)),
      S0 = rnorm(1, 25, 1),
      R0 = rnorm(1, .1, 0.01)
    )
)
fish_mod
# saveRDS(fish_mod, "other/output/fish_jags_mod_030622k_logit_all_converged.rds")
```
# Plot model output - logit, 030622k
```{r}
fish_mod <- readRDS(here(
  "other",
  "output",
  "fish_jags_mod_030622k_logit_all_converged.rds"
))

dev.new()
plot(fish_mod)
thetas <- tidybayes::gather_draws(fish_mod, p[i]) %>%
  group_by(i) %>%
  summarise(
    theta_mean = mean(.value),
    theta_q2.5 = quantile(.value, .025),
    theta_q97.5 = quantile(.value, .975),
    theta_q10 = quantile(.value, .1),
    theta_q90 = quantile(.value, .9)
  ) %>%
  mutate(year = c(2012:2017, 2019, 2012:2019),
         species = c(rep("American Shad", 7), rep("Herring spp.", 8))) %>%
  left_join(raw_sum)

omegas <- tidybayes::gather_draws(fish_mod, mu_p[i]) %>%
  group_by(i) %>%
  summarise(
    omega_mean = mean(.value),
    omega_q2.5 = quantile(.value, .025),
    omega_q97.5 = quantile(.value, .975),
    omega_q10 = quantile(.value, .1),
    omega_q90 = quantile(.value, .9)
  ) %>%
  mutate(species = c("American Shad", "Herring spp."))


thetas %>%
  ggplot() +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_mean),
    size = 0.5,
    color = "red"
  ) +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_q2.5),
    size = 0.5,
    color = "red",
    lty = 2
  ) +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_q97.5),
    size = 0.5,
    color = "red",
    lty = 2
  ) +
  geom_errorbar(aes(xmin = theta_q2.5,
                    xmax = theta_q97.5,
                    y = year), width = 0) +
  geom_errorbar(aes(xmin = theta_q10,
                    xmax = theta_q90,
                    y = year),
                size = 1,
                width = 0) +
  geom_point(aes(x = theta_mean,
                 y = year), size = 4) +
  facet_wrap( ~ species) +
  scale_x_continuous(limits = c(0, 1.1),
                     breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = 2012:2019) +
  theme_bw() +
  theme(text = element_text(size = 14)) +
  labs(y = "", x = "Passage rate") +
  geom_point(aes(x = prop, y = year),
             shape = "x",
             color = "red",
             size = 4) +
  geom_text(aes(x = 1, y = year, label = trials.x),
            nudge_x = .04,
            color = "darkgray") +
  annotate(
    "text",
    x = .94,
    y = 2019,
    label = "n =",
    color = "darkgray"
  )

ggsave(
  "other/figures/jags_nofishhypers_logit_allconverged_run030622k.png",
  width = 7,
  heigh = 5,
  dpi = 400
)
```


# Video time series 
```{r}
# Match up year time series with 
fish_mod <- readRDS(here("other",
                         "output",
                     "fish_jags_mod_030622k_logit_all_converged.rds"
                         ))

# need to load this previous model as fallback parameter wasn't saved in other model
fish_mod2 <- readRDS("other/output/fish_jags_mod_030622m_beta33_nofishhypers_all_converged.rds")

fallback_vector <- tidybayes::gather_draws(fish_mod2, fallback_p_cor)$.value

# collect all theta posterior draws and add fish/year info
theta_draws <- tidybayes::gather_draws(fish_mod, p[i]) %>%
  left_join(data.frame(i = 1:15, 
                       year = c(2012:2017,2019,2012:2019))) %>%
  ungroup() %>%
  mutate(species = c(rep("American Shad", 21000), 
                     rep("River Herring spp.", 24000)),
         fallback = rep(fallback_vector, 15),
         pass_cor = .value/fallback,
         pass_cor2 = .value/.47) 

boxplot(theta_draws$pass_cor~theta_draws$i)
boxplot(theta_draws$pass_cor2~theta_draws$i)
boxplot(theta_draws$.value~theta_draws$i)

# collect omega (fish average) passage rate draws & add fish/year
# (just Shad in 2018 as that is the year with no passage info)
omega_draws_temp <- tidybayes::gather_draws(fish_mod, mu_p[i]) %>%
  filter(i == 1)

omega_draws_temp_shad <- mutate(omega_draws_temp, 
                      year = shad_hist$year[1]) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[2])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[3])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[4])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[5])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[6])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[7])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[8])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[9])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[10])) %>%
  bind_rows(mutate(omega_draws_temp, year = 2018)) %>%
  mutate(species = "American Shad",
         fallback = rep(fallback_vector, 11),
         pass_cor = .value/fallback,
         pass_cor2 = .value/.47)

omega_draws_temp_herring <- mutate(omega_draws_temp, 
                      year = shad_hist$year[1]) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[2])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[3])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[4])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[5])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[6])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[7])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[8])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[9])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[10])) %>%
  mutate(species = "River Herring spp.",
         fallback = rep(fallback_vector, 10),
         pass_cor = .value/fallback,
         pass_cor2 = .value/.47)

omega_draws <- omega_draws_temp_shad %>%
  bind_rows(omega_draws_temp_herring)

rm(omega_draws_temp, 
   omega_draws_temp_shad, 
   omega_draws_temp_herring)

# join posterior draws to fish counts
fish_n_draws <- theta_draws %>%
  # add in global average (omega) passage draws for shad
  # in 2018 with no passage data  
  bind_rows(omega_draws) %>%
  # join with video count data
  left_join(v) %>%
  filter(is.na(n) != TRUE) %>% # remove years with no counts
  arrange(species, year) %>%
  mutate(pass_cor_trunc = ifelse(pass_cor >=1, 1, pass_cor),
         pass_cor2_trunc = ifelse(pass_cor2 >= 1, 1, pass_cor2))

# create corrected n and summarize draws
fish_n <- fish_n_draws %>%
  mutate(nhat = n / pass_cor_trunc) %>%
  group_by(species, year) %>%
  summarise(nhat_med = median(nhat),
            nhat_q2.5 = quantile(nhat, .025),
            nhat_q97.5 = quantile(nhat, .975),
            nhat_q10 = quantile(nhat, .1),
            nhat_q90 = quantile(nhat, .9),
            n = mean(n),
            .groups = "keep") %>%
  bind_rows(data.frame(species = c("American Shad",
                                    "River Herring spp."),
                        year = rep(c(2000, 2004, 2006:2012),2),
                        nhat_med = NA,
                        nhat_q2.5 = NA,
                        nhat_q97.5 = NA,
                        nhat_q10 = NA,
                        nhat_q90 = NA,
                        n = NA)) %>%
  arrange(species, year)

# plot nhat by species and year with credible intervals
fish_n %>%
  filter(species == "American Shad",
         year >= 2013) %>%
  ggplot() +
  geom_ribbon(aes(x = year, ymin = nhat_q2.5+1,
                  ymax = nhat_q97.5+1),
              alpha = 0.25, fill = "steelblue") +
  geom_ribbon(aes(x = year, ymin = nhat_q10,
                  ymax = nhat_q90),
              alpha = 0.25, fill = "steelblue") +
  geom_line(aes(x = year, y = nhat_med+1),
            color = "steelblue", size = .5) +
  geom_line(aes(x = year, y = n+1), size = 0.5, lty = 2,
            color = "steelblue") +
  geom_line(aes(x = year, y = n+1), size = 0.5, lty = 2,
            color = "darkgreen", 
            data = filter(fish_n, 
                          species == "River Herring spp.",
                          year >= 2013)) +
  geom_point(aes(x = year, y = nhat_med+1),
             color = "steelblue", size = 3) +
  geom_errorbar(aes(x = year, ymin = nhat_q2.5+1, 
                  ymax = nhat_q97.5+1),
              alpha = 1, color = "steelblue",
              width = 0) +
  geom_errorbar(aes(x = year, ymin = nhat_q10+1, 
                  ymax = nhat_q90+1),
              alpha = 1, color = "steelblue",
              width = 0, size = 1.5) +
  # facet_wrap(~species) +
  scale_y_log10() +
  scale_x_continuous(breaks = c(2013:2019)) +
  theme_bw() +  
  theme(text = element_text(size = 14)) +
  theme(strip.background = element_rect(fill="white")) +
  labs(x = "", y = "Run size (n)") +
  annotate(geom = "text", x = 2013, y = 225, 
           label = "Shad", color = "steelblue", 
           alpha = 0.75, hjust = 0, size = 4) +
  annotate(geom = "text", x = 2013, y = 30, 
           label = "Herring", color = "darkgreen", 
           alpha = 0.75, hjust = 0, size = 4)
ggsave("other/figures/run_size_passage_cor_uncertainfallback_run030622k_logit.png",
       height = 4, width = 7,
       dpi = 400)

```
# Attraction, staging, passage times
Warning: data are wrong at the moment until antennas are sorted out.
```{r}
fish_sum_plot = fish_sum %>%
  mutate(sp_cat = case_when(species %in% 
                              c("Blueback Herring", 
                                "River Herring",
                                "Alewife") ~
                              "River Herring spp.",
                            species == "American Shad" ~
                              "American Shad",
                            TRUE ~ "NA")) %>%
  filter(sp_cat != "NA") %>%
  select(sp_cat, ends_with("days")) %>%
  pivot_longer(cols = ends_with("days")) %>%
  filter(is.na(value) == FALSE) %>%
  mutate(value = abs(value), # this addresses one -0.0003 time 
         order = case_when(name == "attraction.days" ~ 1,
                           name == "staging.days" ~ 2,
                           name == "passage.days" ~ 4,
                           TRUE ~ 3),
         name = forcats::fct_reorder(name, order),
         metric = case_when(name == "attraction.days" ~ 
                              "Attraction time",
                           name == "staging.days" ~ 
                             "Staging time",
                           name == "passage.days" ~ 
                             "Total passage time",
                           name == "transit.days" ~ 
                             "Transit time"),
         value2 = case_when(name == "transit.days" ~ log10(value),
                            name == "staging.days" ~ log10(value),
                            name == "passage.days" ~ log10(value),
                            TRUE ~ value),
         metric = forcats::fct_reorder(metric, order)) 

fish_sum_plot %>%
  filter(value < 300) %>% # sp_cat == "American Shad"
ggplot() +
  geom_violin(aes(x = sp_cat, # or geom_boxplot
                  y = value+.001, 
                  fill = sp_cat),
              draw_quantiles = c(0.5),
              fill = "transparent",
              color = "darkgray") +
  geom_point(aes(x = sp_cat, y = value+0.001, color = sp_cat),
             size = 3, alpha = .25) +
  facet_wrap(~metric, scales = "free") +
  theme_bw() + 
  labs(x = "", y = "Days", fill = "") +
  scale_y_log10(breaks = c(0.01,0.1, 1, 10)) +
  theme(text = element_text(size = 14),
        # axis.text.x = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = "white")) +
  scale_fill_manual(values = c("steelblue", "darkgreen")) +
  scale_color_manual(values = c("steelblue", "darkgreen"))
  

ggsave("other/figures/passage_time_violin.png",
       height = 4, width = 7,
       dpi = 400)
```
# Summarizing passage metrics
```{r}

# how many total detection
table(fish_sum$species) 
# 26 shad, 6 herring
table(tags$Common.Name) 
# 163 herring = 37 alewife, 50 river herring, 76 blueback
# 51 shad

# shad attract
fish_sum_attract_shad <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(attraction.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, attraction.days)

summary(fish_sum_attract_shad$attraction.days) #
 #      Min.     1st Qu.      Median       Mean    3rd Qu.  Max. 
 # 0.12002639  1.38954873  5.43951  5.397677  8.126332 15.94857870 
sd(fish_sum_attract_shad$attraction.days) # 4.28
length(fish_sum_attract_shad$attraction.days) # 25

# shad stage
fish_sum_stage_shad <- fish_sum %>%
  filter(#attraction.days < 300,
         is.na(staging.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, staging.days)

summary(fish_sum_stage_shad$staging.days) #
 #       Min.     1st Qu.      Median        Mean 3rd Qu.      Max. 
 # 0.00000000  0.01842419  0.41686065  1.67033059  1.97535 19.56337
sd(fish_sum_stage_shad$staging.days) # 3.916
length(fish_sum_stage_shad$staging.days) # 25

# shad transit
fish_sum_transit_shad <- fish_sum %>%
  filter(#attraction.days < 300,
         is.na(transit.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, transit.days)

summary(fish_sum_transit_shad$transit.days) #
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.005   0.007   0.009   0.893   0.048  16.504 
sd(fish_sum_transit_shad$transit.days) # 3.681
length(fish_sum_transit_shad$transit.days) # 20

# shad passage
fish_sum_passage_shad <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(passage.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, passage.days)

summary(fish_sum_passage_shad$passage.days) #
  #    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.373   5.299   6.824   8.569   9.233  27.583
sd(fish_sum_passage_shad$passage.days) # 6.68
length(fish_sum_passage_shad$passage.days) # 20

### Herring

# herring attract
fish_sum_attract_herring <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(attraction.days) == F,
         grepl(species, pattern = "Herring")) %>%
  select(species, attraction.days)

summary(fish_sum_attract_herring$attraction.days) #
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.041   5.119  10.198   8.057  12.064  13.931 
sd(fish_sum_attract_herring$attraction.days) # 7.189
length(fish_sum_attract_herring$attraction.days) # 3

# herring stage
fish_sum_stage_herring <- fish_sum %>%
  filter(#attraction.days < 300,
    is.na(staging.days) == F,
    grepl(species, pattern = "Herring")) %>%
  select(species, staging.days)

summary(fish_sum_stage_herring$staging.days) #
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.000   0.016   2.437   4.353   8.137  11.997 
sd(fish_sum_stage_herring$staging.days) # 5.271
length(fish_sum_stage_herring$staging.days) # 6

# herring transit
fish_sum_transit_herring <- fish_sum %>%
  filter(#attraction.days < 300,
    is.na(transit.days) == F,
    grepl(species, pattern = "Herring")) %>%
  select(species, transit.days)

summary(fish_sum_transit_herring$transit.days) #
#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
# -0.000346  0.004555  0.009456  0.006658  0.010160  0.010865 
sd(fish_sum_transit_herring$transit.days) # 0.006107
length(fish_sum_transit_herring$transit.days) # 3

# herring passage
fish_sum_passage_herring <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(passage.days) == F,
         grepl(species, pattern = "Herring")) %>%
  select(species, passage.days)

summary(fish_sum_passage_herring$passage.days) #
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.04    3.79    7.53    7.53   11.28   15.02 
sd(fish_sum_passage_herring$passage.days) # 10.59
length(fish_sum_passage_herring$passage.days) # 2
  

# look at the 3 individuals that took > 1 year to return
fish_sum_sum_y2 <- fish_sum %>%
  filter(attraction.days >= 300,
         grepl(species, pattern = "Herring|American Shad"))


```