---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---
# Packages
Load packages, functions, and tagging data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(wesanderson)
library(lubridate)
source("scripts/plot_by_tag_function.R")
source("scripts/plot_by_antenna_function.R")
source("scripts/fishsum_function.R")
`%notin%` <- Negate(`%in%`)

# read in tagging data
tags <- read.csv("other/tagging_data_11_2_21.csv") %>%
  filter(Tag.Number != "No Tag Number",
         Tag.Number != "-") %>%
  mutate(tag_short = str_sub(Tag.Number, - 9, - 1),
         tag_time = mdy_hm(Date.and.Time, tz = "America/New_York"))
```
# Compile receiver data
Data are compiled, filtered, and cleaned within these scripts. Anthony should review to ensure all filtering was OK.
```{r}
source("scripts/compile_pings_2012.R")
source("scripts/compile_pings_2013.R")
source("scripts/compile_pings_2014.R")
source("scripts/compile_pings_2015.R")
source("scripts/compile_pings_2016.R")
source("scripts/compile_pings_2017.R")
source("scripts/compile_pings_2018.R")
source("scripts/compile_pings_2019.R")
```
# Export cleaned, compiled data for inspection
```{r}
write.csv(data2012, 
          "output/2012_clean_detection_data.csv", row.names=F)
write.csv(data2013, 
          "output/2013_clean_detection_data.csv", row.names=F)
write.csv(data2014, 
          "output/2014_clean_detection_data.csv", row.names=F)
write.csv(data2015, 
          "output/2015_clean_detection_data.csv", row.names=F)
write.csv(data2016, 
          "output/2016_clean_detection_data.csv", row.names=F)
write.csv(data2017, 
          "output/2017_clean_detection_data.csv", row.names=F)
write.csv(data2018, 
          "output/2018_clean_detection_data.csv", row.names=F)
write.csv(data2019, 
          "output/2019_clean_detection_data.csv", row.names=F)
```
# Plot pings by tag to identify which belong to fish
```{r}
# plot pings by tag number to check and identify fish
plot_by_tag(data2012)
plot_by_tag(data2013)
plot_by_tag(data2014)
plot_by_tag(data2015)
plot_by_tag(data2016)
plot_by_tag(data2017)
plot_by_tag(data2018)
plot_by_tag(data2019)
```
# Plot pings by tag to identify which belong to fish
```{r}
plot_by_antenna(data2012)
plot_by_antenna(data2013)
plot_by_antenna(data2014)
plot_by_antenna(data2015)
plot_by_antenna(data2016)
plot_by_antenna(data2017)
plot_by_antenna(data2018)
plot_by_antenna(data2019)
```
# Extracting & summarizing fish data
extract and summarize data from fish tags
```{r}
fish2012_sum <- fishsum(data2012)
fish2013_sum <- fishsum(data2013)
fish2014_sum <- fishsum(data2014)
fish2015_sum <- fishsum(data2015)
fish2016_sum <- fishsum(data2016)
fish2017_sum <- fishsum(data2017)
fish2018_sum <- fishsum(data2018)
fish2019_sum <- fishsum(data2019)
fish_sum <- fish2019_sum %>%
  bind_rows(fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum) %>%
  arrange(year, Common.Name, tag_time)

# write fish summary csv files for review
# write.csv(fish_sum, "output/fish_sum.csv", row.names = F)

# remove unneeded objects
rm(fish2019_sum, fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum)
```
# Format fish data for "passage proportion" analysis
```{r}
# subset target species and create 1/0 passage variable (did the fish reach A4)
f <- tags %>%
  select(tag_short, Common.Name, tag_time) %>%
  filter(Common.Name %in% c("American Shad",
                            "River Herring",
                            "Blueback Herring")) %>%
  left_join(select(fish_sum, year, tag_short, first_ping_A4), by = "tag_short") %>%
  filter(!duplicated(tag_short)) %>% # remove 3 duplicates (inter-year returns)
  mutate(passage = case_when(is.na(first_ping_A4) ~ 0,
                               TRUE ~ 1),
         year = as.factor(case_when(is.na(year) ~ year(tag_time),
                          TRUE ~ year)),
         year_num = as.numeric(as.factor(year)),
         year_char = paste0("y", year),
         sp = as.factor(case_when(grepl(Common.Name, pattern = "Herring") ~
                          "Herring",
                        TRUE ~ "Shad")))

shad_df <- f %>%
  filter(Common.Name == "American Shad")

herring_df <- f %>%
  filter(grepl(Common.Name, pattern = "Herring"))

cbind.data.frame(
  trials = aggregate(shad_df$passage, 
                     by = list(shad_df$year), length),
  successes = aggregate(shad_df$passage, 
                        by = list(shad_df$year), sum)[,2],
  prop = aggregate(shad_df$passage, by = list(shad_df$year), mean)[,2]
  )

shad_data <- list(
  D = 1,
  N = nrow(filter(f, Common.Name == "American Shad")),
  L = length(unique(filter(f, Common.Name == "American Shad")$year)),
  y = filter(f, Common.Name == "American Shad")$passage,
  ll = as.numeric(as.factor(filter(f, Common.Name == "American Shad")$year)),
  x = rbind(rep(1, nrow(filter(f, Common.Name == "American Shad"))))
)


```
# Defining the Stan model
Modified from here:
https://mc-stan.org/docs/2_28/stan-users-guide/hierarchical-logistic-regression.html
```{r}
cat(file = "scripts/the_stan_model.stan", 
"
data {
  int<lower=1> D;
  int<lower=0> N;
  int<lower=1> L;
  array[N] int<lower=0, upper=1> y;
  array[N] int<lower=1, upper=L> ll;
  array[N] row_vector[D] x;
}
parameters {
  array[D] real mu;
  array[D] real<lower=0> sigma;
  array[L] vector[D] beta;
}
model {
  for (d in 1:D) {
    mu[d] ~ normal(0, 100);
    for (l in 1:L) {
      beta[l, d] ~ normal(mu[d], sigma[d]);
    }
  }
  for (n in 1:N) {
    y[n] ~ bernoulli(inv_logit(x[n] * beta[ll[n]]));
  }
}
"
)
```

```{r}

library(rstan)
rstan_options(auto_write = TRUE)

stan(file = "scripts/the_stan_model.stan",
     model_name = "shad passage",
     data = shad_data)
```
# Trying with brms to start
https://bookdown.org/ajkurz/DBDA_recoded/hierarchical-models.html#doing-it-with-jags-brms.
```{r}
library(brms)
shad_fit <-
  brm(data = shad_df,
      family = bernoulli(link = logit),
      passage ~ 1 + (1 | year_char),
      prior = c(prior(normal(0, 1.5), class = Intercept),
                prior(normal(0, 1), class = sd)),
      iter = 160000, warmup = 80000, thin = 100, 
      chains = 3, cores = 3,
      seed = 9)

# saveRDS(shad_fit, "other/shad_fit.rds")
shad_fit <- readRDS("other/shad_fit.rds")
plot(shad_fit)

# convert posterior draws to df
shad_post <- as_draws_df(shad_fit, add_chain = T) %>%
    as.data.frame() %>%
    select(1:9) %>%
    rename(mean = 1, rand_sd = 2, y12 = 3, y13 = 4, 
           y14 = 5, y15 = 6, y16 = 7, y17 = 8, y19 = 9)

hist(plogis(shad_post$y14))

print(shad_fit)

shad_plot <- cbind.data.frame(
  mean = apply(shad_post, 2, mean),
  q25 = apply(shad_post, 2, function(x) quantile(x, 0.5)),
  q75 = apply(shad_post, 2, function(x) quantile(x, 0.75)),
  q2.5 = apply(shad_post, 2, function(x) quantile(x, 0.025)),
  q97.5 = apply(shad_post, 2, function(x) quantile(x, 0.975)),
  q10 = apply(shad_post, 2, function(x) quantile(x, 0.1)),
  q90 = apply(shad_post, 2, function(x) quantile(x, 0.9))) %>%
  mutate(year = c(NA, NA, 2012:2017, 2019),
         names = colnames(shad_post)) 

shad_plot %>%
  filter(is.na(year)==F) %>%
  ggplot() +
  geom_vline(xintercept = plogis(shad_plot[1,]$mean), size = 0.5,
             color = "red") +
  geom_vline(xintercept = plogis(shad_plot[1,]$q2.5), size = 0.5,
             color = "red", lty = 2) +
  geom_vline(xintercept = plogis(shad_plot[1,]$q97.5), size = 0.5,
             color = "red", lty = 2) +
  geom_errorbar(aes(xmin = plogis(q2.5), xmax = plogis(q97.5), 
                    y = year), width = 0) +
  geom_errorbar(aes(xmin = plogis(q10), xmax = plogis(q90), 
                    y = year), width = 0, size = 1) +
  geom_point(aes(x = plogis(mean), y = year), size = 4,
             color = "black") +
  scale_x_continuous(limits = c(0,1)) +
  theme_bw() +
  theme(text = element_text(size = 14)) +
  labs(y = "", x = "Passage rate", title = "American Shad")

# ggsave("figures/shad_passage_placeholder.png", height = 6, width = 6, dpi = 400)
  
```

```{r}


```

