---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---

```{r}
# Packages
library(dplyr)
library(ggplot2)
library(stringr)
library(wesanderson)
library(lubridate)
```
# Compile receiver data from 2017
Type: D (data), E (information), or B (bad)
Date: MM/DD/YYYY
Time: HH:MM:SS.SS
Unknown1: HH:MM:SS.SS?
Unknown2: HA, HR, or HW
Tag: 27 long values
Antenna: A1, A3, A4 (and "1")
Unknown3: 1-25, 27-29, 31, 34-36, 38, 4454
Unknown4: 871 values
```{r}
# List of files to read
datadir <- "data/receiver_files/2017"
files <- list.files(datadir)

# create empty list to contain formated data
file_data_list <- list()

# Loop through files and format data
for(i in 1:length(files)){
  # Read data
  filename <- files[i]
  print(paste(i, filename))
  filepath <- paste(datadir, filename, sep="/")
  data_all <- read.table(filepath, col.names=c("type", "date", "time",
                                               "unknown1", "unknown2", 
                                               "tag", "antenna", 
                                               "unknown3", "unknown4",
                                           paste("extra", 1:5, sep="")),
                         fill=T, as.is=T, na.strings="") %>%
  # Type: D (data) or E (error/info) or B (bad - only one file)
    filter(type == "D")

  # Only do the following if there is data
  if(nrow(data_all)>0){
  
    # Up or down file?
    data <- data_all %>%
      mutate(direction = case_when(grepl("up|Up|UP", 
                                         filename) ~ "up",
                                   grepl("down|Down|DOWN", 
                                         filename) ~ "down",
                                   TRUE ~ "direction unknown"),
             # flag antenna names that don't match expectations
             flag = case_when(direction == "up" & 
                                antenna %in% c("A1", "A2") ~ "mismatch",
                              direction == "down" & 
                                antenna %in% c("A3", "A4") ~ "mismatch",
                              TRUE ~ "OK"),
             file = filename)
    
    # Rename antennas if necessary (old - check on this)
    # Up: A1 becomes A4; A3 stays the same
    # Down: A3 becomes A2; A1 stays the same
    # if(direction=="up"){data$antenna[data$antenna=="A1"] <- "A4"}
    # if(direction=="down"){data$antenna[data$antenna=="A3"] <- "A2"}

   # Merge datasets
    file_data_list[[i]] <- data
    
  }
  
}

# set number of digits for seconds
options(digits.secs=2)
options(digits = 11)

# bind all data files together and format them
data2017 <- do.call(rbind, file_data_list) %>%
  select(file, date, time, antenna, tag, direction, flag) %>%
  # correct 2 reading errors manually from these three lines
  # 6_26_17 Down.TXT file at 2017-06-26 09:36:39.71
  # 6_26_17 Down.TXT file at 2017-06-25 21:26:28.43
  # 6_26_17 Down.TXT file at 2017-06-24 19:11:43.60
  mutate(tag = case_when(tag == "1" ~ "000000180843403", 
                         tag == "A1" ~ "0000_0000000180843403",
                         tag == "00šúÁ000000180843440" ~ 
                           "0000_0000000180843440",
                         TRUE ~ tag),
         antenna = case_when(antenna %in% c("2", "4440") ~ "A1", 
                         TRUE ~ antenna)) %>%
  # delete glitch row
  # 5_15_17 Down 2.TXT 2019-03-02 17:05:43.25
  filter(antenna != "0000") %>%
  # remove one weird ping from 2018-06-02
  filter(date != "2018-06-02") %>%  
  # make shorter tag number
  mutate(tag_short = str_sub(tag, - 9, - 1)) %>%
  # format date and time
  mutate(date = as.Date(date),
         # datetime = strptime(paste(date, time), 
         #                     format="%Y-%m-%d %H:%M:%OS"),
         datetime = ymd_hms(paste(date, time), tz = "America/New_York"))
  

# # saving this in case I need it later
#   mutate(year = year(datetime2),
#          ord = yday(datetime2),
#          hour = hour(datetime2),
#          min = minute(datetime2),
#          sec = second(datetime2),
#          dhour = year + (ord + (hour + (min + sec/60)/60)/24)/
#          )

# check for antenna naming issues
unique(data2017$flag)
unique(data2017$tag)
unique(data2017$tag_short)

# subset final columns
data2017 <- data2017 %>%
  select(file, date, time, datetime, antenna, tag, tag_short) %>%
  arrange(antenna, tag, datetime)

# rm(file_data_list)

# read in tagging data
tags <- read.csv("other/tagging_data_11_2_21.csv") %>%
  filter(Tag.Number != "No Tag Number",
         Tag.Number != "-") %>%
  mutate(tag_short = str_sub(Tag.Number, - 9, - 1),
         tag_time = mdy_hm(Date.and.Time, tz = "America/New_York"))

```
# Export 2017 data for inspection
```{r}
write.csv(data2017, "output/2017_clean_detection_data.csv", row.names=F)
```
# Plot pings by tag to identify which belong to fish
```{r}
data2017 %>%
  ggplot() +
  geom_point(aes(x = datetime, y = as.factor(tag_short), color = antenna),
             alpha = .8, pch = 1) +
  scale_color_manual(values = c(wes_palettes$FantasticFox1[1:3],
                                "firebrick")) +
  theme_bw() +
  labs(x = "", y = "") +
  theme(text = element_text(size = 15))
  
# ggsave("figures/tag_check.png", height = 7, width = 10, dpi = 400)
```

# Plot pings by tag to identify which belong to fish
```{r}
data2017 %>%
  ggplot() +
  geom_point(aes(x = datetime, y = as.factor(antenna)),
             alpha = .5, pch = 1) +
  theme_bw() +
  labs(x = "", y = "") +
  theme(text = element_text(size = 15))
  
# ggsave("figures/antenna_check.png", height = 7, width = 10, dpi = 400)
```

# Extracting & summarizing fish
```{r}
# extract data only from fish tags
fish2017 <- data2017 %>%
  filter(tag_short %in% c("181177668", "000093631",
                          "000093630", "000093628",
                          "000093624", "000093600", "000093602",
                          "000093603", "000093604", 
                          "000093605", "000093606",
                          "000008917", "000008918")) %>%
  arrange(tag_short, datetime)

# summarize data by antenna and fish
fish2017_sum <- fish2017 %>%
  group_by(tag_short, antenna) %>%
  summarize(
    first_ping = min(datetime),
    last_ping = max(datetime),
    n = length(datetime),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(names_from = antenna,
                     values_from = c(first_ping, last_ping, n)) %>%
  # add in tagging time and common name
  left_join(tags) %>%
  select(tag_short, Common.Name, tag_time, 2:13) %>%
  # calculate durations
  # attraction time (tag_time -> min A1)
  # staging time (min A1 -> max A1)
  # transit time (max A1 -> max A4)
  # passage time (sum of other 3)
  mutate(
    attraction.days = as.numeric(as.duration(
      interval(start = tag_time,
               end = first_ping_A1)
    )) / (3600 * 24),
    staging.days =
      as.numeric(as.duration(
        interval(start = first_ping_A1,
                 end = last_ping_A1)
      )) / (3600 * 24),
    transition.days = as.numeric(as.duration(
      interval(start = last_ping_A1,
               end = last_ping_A4)
    )) / (3600 * 24),
    passage.days = as.numeric(as.duration(
      interval(start = tag_time,
               end = last_ping_A4)
    )) / (3600 * 24)
  )

# write csv for review
write.csv(fish2017_sum, "output/fish2017_sum.csv", row.names = F)

```
# calculate times
attraction time (tag_time -> min A1)
staging time (min A1 -> max A1)
transit time (max A1 -> max A4)
passage time (sum of other 3)
```{r}


fish2017_sum <- fish2017_sum %>%
  mutate(attraction = as.duration(interval(start = tag_time,
                                          end = first_ping_A1)))


```

