---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---

# Packages
Load packages, functions, & data; format data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(wesanderson)
library(lubridate)
library(readxl)
source("scripts/fishsum_function.R")
`%notin%` <- Negate(`%in%`)

# read in tagging data
# note: currently excludes herring w/ "226000011055/35/85" tag

tags <- read_xlsx(
  "other/1 IFW Tagging Data Golden Master 2_25_22.xlsx",
  sheet = "All Tagged Fish") %>%
  rename(tag = 'Tag Number',
         datetime = as.character('Date and Time'),
         Common.Name = 'Common Name') %>%
  mutate(tag = as.numeric(tag)) %>%
  filter(is.na(tag) == FALSE) %>%
  mutate(tag_short = str_sub(tag, - 9, - 1),
         tag_time = ymd_hms(datetime, 
                           tz = "America/New_York"))

# video time series - new
v_time <- 
  read_xlsx("other/IFW Video Data Golden Master_4_24_22.xlsx",
               sheet = "Grouped") %>%
  select(Date, time = Time, species = Species) %>%
  mutate(year = year(Date),
         # note: time read in as decimal hours (a prop. of 24 hr)
         time = as.numeric(as.character(time))*24, 
         hour = trunc(time),
         daytime = case_when(time/24 > (7/24) & 
                               time/24 < (19/24) ~ 1,
                             TRUE ~ 0)) %>%
  group_by(species, year, daytime) %>%
  tally() %>%
  filter(species %in% c("American Shad", "River Herring spp.")) %>%
  arrange(species, year, daytime) %>%
  group_by(species, year) %>%
  mutate(pday = n/sum(n))
# write.csv(v_time, here("output", 
#                           "proportion_counts_in_daytime.csv"), 
#           row.names=F)

v_temp <- v_time %>%
  group_by(species, year) %>%
  summarise(n = sum(n))

# historic Island Farm Wier Video counts
shad_hist <- data.frame(species = "American Shad",
                        year = c(1996:2005),
                        n = c(48, 76, 149, 582, NA, 717, 
                                1114, 362, NA, 13))

herring_hist <- data.frame(species = "River Herring spp.",
                           year = c(1996:2005),
                        n = c(0, 117, 9, 114, NA, 61, 
                                17, 12, NA, 0))

# combine recent and historic video count data
v <- v_temp %>%
  bind_rows(shad_hist, herring_hist) %>%
  arrange(species, year)

rm(v_temp)

# Compile receiver data
# Data are compiled, filtered, and cleaned within these scripts. Anthony should review to ensure all filtering was OK.
source("scripts/compile_pings_2012.R")
source("scripts/compile_pings_2013.R")
source("scripts/compile_pings_2014.R")
source("scripts/compile_pings_2015.R")
source("scripts/compile_pings_2016.R")
source("scripts/compile_pings_2017.R")
source("scripts/compile_pings_2018.R")
source("scripts/compile_pings_2019.R")

# Extracting & summarizing fish tag data
fish2012_sum <- fishsum(data2012)
fish2013_sum <- fishsum(data2013)
fish2014_sum <- fishsum(data2014)
fish2015_sum <- fishsum(data2015)
fish2016_sum <- fishsum(data2016)
fish2017_sum <- fishsum(data2017)
fish2018_sum <- fishsum(data2018)
fish2019_sum <- fishsum(data2019)
fish_sum <- fish2019_sum %>%
  bind_rows(fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum) %>%
  arrange(species, tag_time)

# remove unneeded objects
rm(fish2019_sum, fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum)

# subset target species and create 1/0 passage variable (did the fish reach A4)
f <- tags %>%
  select(tag_short, Common.Name, tag_time) %>%
  filter(Common.Name %in% c("American Shad",
                            "River Herring",
                            "Blueback Herring",
                            "Alewife")) %>%
  left_join(select(fish_sum, year = detect_year, tag_short, passage.days), by = "tag_short") %>%
  # remove 1 herring tagged in 2014 but only detected in 2015
  filter(is.na(passage.days) | passage.days < 300) %>%
  mutate(passage = case_when(is.na(passage.days) ~ 0,
                               TRUE ~ 1),
         year = as.factor(case_when(is.na(year) ~ year(tag_time),
                          TRUE ~ year)),
         year_num = as.numeric(as.factor(year)),
         year_char = paste0("y", year),
         sp = as.factor(case_when(grepl(Common.Name, 
                                        pattern="Herring|Alewife")~
                          "Herring",
                        TRUE ~ "Shad")))

shad_df <- f %>%
  filter(sp == "Shad")

herring_df <- f %>%
  filter(sp == "Herring")

# examine shad data
shad_raw_sum <- cbind.data.frame(
  trials = aggregate(shad_df$passage, 
                     by = list(shad_df$year), length),
  successes = aggregate(shad_df$passage, 
                        by = list(shad_df$year), sum)[,2],
  prop = aggregate(shad_df$passage, by = list(shad_df$year), mean)[,2]
  ) %>%
  mutate(year = as.numeric(as.character(trials.Group.1)),
         species = "American Shad") %>%
  select(-trials.Group.1)

# examine herring data
herring_raw_sum <- cbind.data.frame(
  trials = aggregate(herring_df$passage, 
                     by = list(herring_df$year), length),
  successes = aggregate(herring_df$passage, 
                        by = list(herring_df$year), sum)[,2],
  prop = aggregate(herring_df$passage, by = list(herring_df$year), mean)[,2]
  ) %>%
  mutate(year = as.numeric(as.character(trials.Group.1)),
         species = "Herring spp.") %>%
  select(-trials.Group.1)

raw_sum <- rbind(shad_raw_sum, herring_raw_sum)

shad_data <- list(
  D = 1,
  N = nrow(filter(f, Common.Name == "American Shad")),
  L = length(unique(filter(f, Common.Name == "American Shad")$year)),
  y = filter(f, Common.Name == "American Shad")$passage,
  ll = as.numeric(as.factor(filter(f, Common.Name == "American Shad")$year)),
  x = rbind(rep(1, nrow(filter(f, Common.Name == "American Shad"))))
)

# Format fish data for "passage proportion" analysis

# combine raw data summaries
both <- rbind(shad_raw_sum, herring_raw_sum) %>%
  select(N = trials.x, z = successes, -prop, year) %>%
  mutate(year = as.numeric(as.factor(year)),
         fish = c(rep(1,7), rep(2,8)),
         fishyears = 1:15) 

nfishyears <- 15
nyears <- 8
nfish <- 2

# bundle data for jags
fish_jags_data <- list(
  nfish = nfish,
  nfishyears = nfishyears,
  z = both$z,
  N = both$N,
  c = both$fish,
  fallback_detect = c(rep(1,33), rep(0,106)),
  fallback_n = 139
)
```
# Save data and diagnostic plot files for inspection
Need to be checked by Anthony.
```{r}
# source("scripts/save_diagnostics.R")
```
# JAGS model for fish passage proportion
logit-normal with hyperparameter for each fish
```{r}

cat(file = "scripts/fish_mod_logit.txt",
"
model{
for(s in 1:nfishyears){
  z[s] ~ dbin(ilogit(lp[s]), N[s])
  lp[s] ~ dnorm(mu_logit_p[c[s]], tau_logit_p[c[s]] )
  logit(p[s]) <- lp[s]
}  

for( c in 1:nfish){
mu_logit_p[c] ~ dnorm(0, 0.001)
tau_logit_p[c] ~ dgamma(0.001, 0.001)
logit(mu_p[c]) <- mu_logit_p[c]
}

fallback_p ~ dunif(0,1) # dbeta(3,3)
for(i in 1:length(fallback_detect)){
fallback_detect[i] ~ dbern(fallback_p)
}

# corrected for imperfect detection in the arrays
fallback_p_cor <- fallback_p / 0.5


}
"
)
```
# Run model
```{r}
fish_mod_new <- jagsUI::jags(
  data = fish_jags_data,
  model.file = "scripts/fish_mod_logit.txt",
  n.iter = 1000000,
  n.burnin = 500000,
  parameters.to.save = c(
    "theta",
    "omega",
    "kappa",
    "S0",
    "R0",
    "p",
    "mu_p",
    "mu_logit_p",
    "tau_logit_p",
    "lp",
    "fallback_p",
    "fallback_p_cor"
  ),
  n.chains = 3,
  parallel = T,
  n.thin = 200,
  inits = function()
    list(
      lp = c(rnorm(7, 0, 0.1),
             rnorm(8, -10, 0.1)),
      omega =
        c(rnorm(1, .47, .01),
          rnorm(1, .01, .01)),
      kappaMinusTwo = c(rnorm(1, 400, 2),                                                      rnorm(1, 1000, 2)),
      S0 = rnorm(1, 25, 1),
      R0 = rnorm(1, .1, 0.01)
    )
)
fish_mod_new
# saveRDS(fish_mod_new, "other/output/fish_jags_mod_052022a_logit_all_converged.rds")
```
# Plot model output - logit, 030622k
```{r}
# old version with 2 herring passed
# fish_mod <- readRDS(here(
#   "other",
#   "output",
#   "fish_jags_mod_030622k_logit_all_converged.rds"
# ))

# new version with 1 herring passed
fish_mod <- readRDS(here(
  "other",
  "output",
  "fish_jags_mod_052022a_logit_all_converged.rds"
))

# dev.new()
# plot(fish_mod)
thetas <- tidybayes::gather_draws(fish_mod, p[i]) %>%
  group_by(i) %>%
  summarise(
    theta_mean = mean(.value),
    theta_q2.5 = quantile(.value, .025),
    theta_q97.5 = quantile(.value, .975),
    theta_q10 = quantile(.value, .1),
    theta_q90 = quantile(.value, .9)
  ) %>%
  mutate(year = c(2012:2017, 2019, 2012:2019),
         species = c(rep("American Shad", 7), rep("Herring spp.", 8))) %>%
  left_join(raw_sum) %>%
  mutate(x = c(rep(1.05, 7), rep(.09, 8)),
         n_label = paste0(c(rep("", 6),"n =",
                            rep("", 7),"n ="), trials.x))

omegas <- tidybayes::gather_draws(fish_mod, mu_p[i]) %>%
  group_by(i) %>%
  summarise(
    omega_mean = mean(.value),
    omega_q2.5 = quantile(.value, .025),
    omega_q97.5 = quantile(.value, .975),
    omega_q10 = quantile(.value, .1),
    omega_q90 = quantile(.value, .9)
  ) %>%
  mutate(species = c("American Shad", "Herring spp."))


thetas %>%
  ggplot() +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_mean),
    size = 0.5,
    color = "red"
  ) +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_q2.5),
    size = 0.5,
    color = "red",
    lty = 2
  ) +
  geom_vline(
    data = omegas,
    aes(xintercept = omega_q97.5),
    size = 0.5,
    color = "red",
    lty = 2
  ) +
  geom_errorbar(aes(xmin = theta_q2.5,
                    xmax = theta_q97.5,
                    y = year), width = 0) +
  geom_errorbar(aes(xmin = theta_q10,
                    xmax = theta_q90,
                    y = year),
                size = 1,
                width = 0) +
  geom_point(aes(x = theta_mean,
                 y = year), size = 4) +
  facet_wrap( ~ species, scales = "free") +
  # scale_x_continuous(#limits = c(0, 1.1),
  #                    breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(breaks = 2012:2019) +
  theme_bw() +
  theme(text = element_text(size = 14)) +
  labs(y = "", x = "Passage rate") +
  geom_point(aes(x = prop, y = year),
             shape = "x",
             color = "red",
             size = 4) +
  geom_text(aes(x = x, y = year, 
                label = n_label),
            nudge_x = .02,
            color = "darkgray",
            hjust = 1) 



# ggsave(
#   "other/figures/jags_nofishhypers_logit_allconverged_run052222a.png",
#   width = 7,
#   heigh = 5,
#   dpi = 400
# )
```

# Video time series 
```{r}
# Match up year time series with 
fish_mod <- readRDS(here("other",
                         "output",
                     "fish_jags_mod_030622k_logit_all_converged.rds"
                         ))

# need to load this previous model as fallback parameter wasn't saved in other model
fish_mod2 <- readRDS("other/output/fish_jags_mod_030622m_beta33_nofishhypers_all_converged.rds")

fallback_vector <- tidybayes::gather_draws(fish_mod2, fallback_p_cor)$.value

# collect all theta posterior draws and add fish/year info
theta_draws <- tidybayes::gather_draws(fish_mod, p[i]) %>%
  left_join(data.frame(i = 1:15, 
                       year = c(2012:2017,2019,2012:2019))) %>%
  ungroup() %>%
  mutate(species = c(rep("American Shad", 21000), 
                     rep("River Herring spp.", 24000)),
         fallback = rep(fallback_vector, 15),
         pass_cor = case_when(species == "American Shad" ~
                                .value,
                              species == "River Herring spp." ~
                                .value/fallback),
         pass_cor2 = .value/.47) 

boxplot(theta_draws$pass_cor~theta_draws$i)
boxplot(theta_draws$pass_cor2~theta_draws$i)
boxplot(theta_draws$.value~theta_draws$i)

# collect omega (fish average) passage rate draws & add fish/year
# (just 2020 2021, & early years with no passage info)
# (also 2018 missing for Shad)
omega_draws_temp <- tidybayes::gather_draws(fish_mod, mu_p[i]) %>%
  filter(i == 1)

omega_draws_temp_shad <- mutate(omega_draws_temp, 
                      year = shad_hist$year[1]) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[2])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[3])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[4])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[5])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[6])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[7])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[8])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[9])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[10])) %>%
  bind_rows(mutate(omega_draws_temp, year = 2018)) %>%
  bind_rows(mutate(omega_draws_temp, year = 2020)) %>%
  bind_rows(mutate(omega_draws_temp, year = 2021)) %>%
  mutate(species = "American Shad",
         fallback = rep(fallback_vector, 13),
         pass_cor = .value,
         pass_cor2 = .value/.47)

omega_draws_temp <- tidybayes::gather_draws(fish_mod, mu_p[i]) %>%
  filter(i == 2)

omega_draws_temp_herring <- mutate(omega_draws_temp, 
                      year = shad_hist$year[1]) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[2])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[3])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[4])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[5])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[6])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[7])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[8])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[9])) %>%
  bind_rows(mutate(omega_draws_temp, year = shad_hist$year[10])) %>%
  bind_rows(mutate(omega_draws_temp, year = 2020)) %>%
  bind_rows(mutate(omega_draws_temp, year = 2021)) %>%
  mutate(species = "River Herring spp.",
         fallback = rep(fallback_vector, 12),
         pass_cor = .value/fallback,
         pass_cor2 = .value/.47)

omega_draws <- omega_draws_temp_shad %>%
  bind_rows(omega_draws_temp_herring)

rm(omega_draws_temp, 
   omega_draws_temp_shad, 
   omega_draws_temp_herring)

# join posterior draws to fish counts
fish_n_draws <- theta_draws %>%
  # add in global average (omega) passage draws for shad
  # in 2018 with no passage data  
  bind_rows(omega_draws) %>%
  # join with video count data
  left_join(v) %>%
  filter(is.na(n) != TRUE) %>% # remove years with no counts
  arrange(species, year) %>%
  mutate(pass_cor_trunc = ifelse(pass_cor >=1, 1, pass_cor),
         pass_cor2_trunc = ifelse(pass_cor2 >= 1, 1, pass_cor2))

# create corrected n and summarize draws
fish_n <- fish_n_draws %>%
  mutate(nhat = n / pass_cor_trunc) %>%
  group_by(species, year) %>%
  summarise(nhat_med = median(nhat),
            nhat_q2.5 = quantile(nhat, .025),
            nhat_q97.5 = quantile(nhat, .975),
            nhat_q10 = quantile(nhat, .1),
            nhat_q90 = quantile(nhat, .9),
            n = mean(n),
            .groups = "keep") %>%
  bind_rows(data.frame(species = c("American Shad",
                                    "River Herring spp."),
                        year = rep(c(2000, 2004, 2006:2012),2),
                        nhat_med = NA,
                        nhat_q2.5 = NA,
                        nhat_q97.5 = NA,
                        nhat_q10 = NA,
                        nhat_q90 = NA,
                        n = NA)) %>%
  arrange(species, year) %>%
  mutate(nhat_q90 = 
           case_when(species == "River Herring spp." ~
                                nhat_med,
                              TRUE ~ nhat_q90),
         nhat_q97.5 = 
           case_when(species == "River Herring spp." ~
                                nhat_med,
                              TRUE ~ nhat_q97.5))

# plot nhat by species and year with credible intervals
fish_n %>%
  filter(year >= 2013) %>%
  ggplot() +
  geom_ribbon(aes(x = year, ymin = nhat_q2.5,
                  ymax = nhat_q97.5, fill = species),
              alpha = 0.25) +
  geom_ribbon(aes(x = year, ymin = nhat_q10,
                  ymax = nhat_q90, fill = species),
              alpha = 0.25) +
  geom_line(aes(x = year, y = nhat_med, color = species),
            size = .5) +
  geom_line(aes(x = year, y = n, color = species), 
            size = 0.5, lty = 2) +
  geom_point(aes(x = year, y = nhat_med, color = species),
             size = 3) +
  geom_errorbar(aes(x = year, ymin = nhat_q2.5, 
                  ymax = nhat_q97.5, color = species),
              alpha = 1,
              width = 0) +
  geom_errorbar(aes(x = year, ymin = nhat_q10, 
                  ymax = nhat_q90, color = species),
              alpha = 1,
              width = 0, size = 1.5) +
  scale_fill_manual(values = c("steelblue", "steelblue")) +
  scale_color_manual(values = c("steelblue", "steelblue")) +
  facet_wrap(~species) +
  scale_y_log10() +
  guides(fill = "none", color = "none") +
  scale_x_continuous(breaks = c(2013:2021)) +
  theme_bw() +  
  theme(text = element_text(size = 14),
        axis.text.x = element_text(size = 8)) +
  theme(strip.background = element_rect(fill="white")) +
  labs(x = "", y = "Run size (n)")

ggsave("other/figures/run_size_passage_cor_uncertainfallback_run030622k_logit_20220426_newdata2.png",
       height = 4, width = 7,
       dpi = 400)

```
# Attraction, staging, passage times
```{r}
fish_sum_plot = fish_sum %>%
  mutate(sp_cat = case_when(species %in% 
                              c("Blueback Herring", 
                                "River Herring",
                                "Alewife") ~
                              "River Herring spp.",
                            species == "American Shad" ~
                              "American Shad",
                            TRUE ~ "NA")) %>%
  filter(sp_cat != "NA") %>%
  select(sp_cat, ends_with("days")) %>%
  pivot_longer(cols = ends_with("days")) %>%
  filter(is.na(value) == FALSE) %>%
  mutate(value = abs(value), # this addresses one -0.0003 time 
         order = case_when(name == "attraction.days" ~ 1,
                           name == "staging.days" ~ 2,
                           name == "passage.days" ~ 4,
                           TRUE ~ 3),
         name = forcats::fct_reorder(name, order),
         metric = case_when(name == "attraction.days" ~ 
                              "Attraction time",
                           name == "staging.days" ~ 
                             "Staging time",
                           name == "passage.days" ~ 
                             "Total passage time",
                           name == "transit.days" ~ 
                             "Transit time"),
         value2 = case_when(name == "transit.days" ~ log10(value),
                            name == "staging.days" ~ log10(value+0.001),
                            TRUE ~ value),
         metric = forcats::fct_reorder(metric, order))

the_colors <- rep(c("darkgray", "transparent"), 4)

fish_sum_plot %>%
  filter(value < 300,
         !as.logical(name == "transit.days" & 
                       value > .9)) %>% # sp_cat == "American Shad"
ggplot() +
  geom_boxplot(aes(x = sp_cat, # or geom_boxplot
                  y = value, 
                  fill = sp_cat),
              draw_quantiles = c(0.5),
              fill = "transparent",
              colour = the_colors, 
              outlier.color = "transparent") +
  geom_point(aes(x = sp_cat, 
                 y = value, color = sp_cat),
             size = 3, alpha = .25) +
  facet_wrap(~metric, scales = "free") +
  theme_bw() + 
  labs(x = "", y = "Days", fill = "") +
  # scale_y_log10(breaks = c(0.01,0.1, 1, 10)) +
  theme(text = element_text(size = 14),
        # axis.text.x = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = "white")) +
  scale_fill_manual(values = c("steelblue", "darkgreen")) +
  scale_color_manual(values = c("steelblue", "darkgreen"))
  

ggsave("other/figures/passage_time_boxplot4.png",
       height = 4, width = 7,
       dpi = 400)
```
# Summarizing passage metrics
```{r}

# how many total detection
table(fish_sum$species) 
# 26 shad, 5 herring
table(tags$Common.Name) 
# 163 herring = 37 alewife, 50 river herring, 76 blueback
# Anthony edited this in the ms to 167 
# (+ 4 fish found in records, not in tagging database)
# 51 shad

# shad attract
fish_sum_attract_shad <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(attraction.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, attraction.days)

summary(fish_sum_attract_shad$attraction.days) #
 #      Min.     1st Qu.      Median       Mean    3rd Qu.  Max. 
 # 0.12002639  1.38954873  5.43951  5.397677  8.126332 15.94857870 
sd(fish_sum_attract_shad$attraction.days) # 4.28
length(fish_sum_attract_shad$attraction.days) # 25

# shad stage
fish_sum_stage_shad <- fish_sum %>%
  filter(#attraction.days < 300,
         is.na(staging.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, staging.days)

summary(fish_sum_stage_shad$staging.days) #
 #       Min.     1st Qu.      Median        Mean 3rd Qu.      Max. 
 # 0.00000000  0.01842419  0.41686065  1.67033059  1.97535 19.56337
sd(fish_sum_stage_shad$staging.days) # 3.916
length(fish_sum_stage_shad$staging.days) # 25

# shad transit
fish_sum_transit_shad <- fish_sum %>%
  filter(#attraction.days < 300,
         is.na(transit.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, transit.days)

summary(fish_sum_transit_shad$transit.days) #
# Min.  1st Qu. Median   Mean 3rd Qu. Max. 
# 0.0045 0.0072 0.0093 0.0702 0.0477 0.9988
sd(fish_sum_transit_shad$transit.days) # 0.2916
length(fish_sum_transit_shad$transit.days) # 20

# shad passage
fish_sum_passage_shad <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(passage.days) == F,
         grepl(species, pattern = "American Shad")) %>%
  select(species, passage.days)

summary(fish_sum_passage_shad$passage.days) #
  #  Min. 1st Qu. Median  Mean 3rd Qu.  Max. 
  #  0.373  5.299  6.824  7.746  9.233 20.884 
sd(fish_sum_passage_shad$passage.days) # 5.023
length(fish_sum_passage_shad$passage.days) # 20

### Herring

# herring attract
fish_sum_attract_herring <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(attraction.days) == F,
         grepl(species, pattern = "Herring")) %>%
  select(species, attraction.days)

summary(fish_sum_attract_herring$attraction.days) #
#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
# 10.197589 11.131015 12.064440 12.064440 12.997866 13.931292 
sd(fish_sum_attract_herring$attraction.days) # 2.6401264713
length(fish_sum_attract_herring$attraction.days) # 3

# herring stage
fish_sum_stage_herring <- fish_sum %>%
  filter(#attraction.days < 300,
    is.na(staging.days) == F,
    grepl(species, pattern = "Herring")) %>%
  select(species, staging.days)

summary(fish_sum_stage_herring$staging.days) #
 #  Min.    1st Qu.       Median       Mean      3rd Qu.    Max. 
 # 0.0  0.062577083  4.810761343  5.223404468  9.246379745 11.997
sd(fish_sum_stage_herring$staging.days) # 5.3888
length(fish_sum_stage_herring$staging.days) # 5

# herring transit
fish_sum_transit_herring <- fish_sum %>%
  filter(#attraction.days < 300,
    is.na(transit.days) == F,
    grepl(species, pattern = "Herring")) %>%
  select(species, transit.days)

summary(fish_sum_transit_herring$transit.days) #
#  Min.   1st Qu.  Median   Mean   3rd Qu.    Max. 
# 0.0095  0.009808 0.01016 0.01016 0.01051 0.01086
sd(fish_sum_transit_herring$transit.days) # 0.00099625125564
length(fish_sum_transit_herring$transit.days) # 3

# herring passage
fish_sum_passage_herring <- fish_sum %>%
  filter(attraction.days < 300,
         is.na(passage.days) == F,
         grepl(species, pattern = "Herring")) %>%
  select(species, passage.days)

summary(fish_sum_passage_herring$passage.days) #
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.04    3.79    7.53    7.53   11.28   15.02 
sd(fish_sum_passage_herring$passage.days) # 10.59
length(fish_sum_passage_herring$passage.days) # 2
  

# look at the 3 individuals that took > 1 year to return
fish_sum_sum_y2 <- fish_sum %>%
  filter(attraction.days >= 300,
         grepl(species, pattern = "Herring|American Shad"))

```

# Stairstep figure
```{r}

stairs <- fish_sum %>%
  filter(species %in% c("American Shad", "Blueback Herring")) %>%
  mutate(farthest = case_when(!is.na(n_A4) ~ 4,
                              !is.na(n_A3) & is.na(n_A4) ~ 3,
                              !is.na(n_A2) & is.na(n_A3) &
                                is.na(n_A4) ~ 2,
                              is.na(n_A2) & is.na(n_A3) &
                                is.na(n_A4) ~ 1)
                              ) %>%
  right_join(tags) %>%
  filter(grepl(Common.Name, 
               pattern = "Herring|American Shad|Alewife")) %>%
  mutate(sp = case_when(grepl(Common.Name, 
                                   pattern = "Herring|Alewife") ~
                               "River Herring",
                        grepl(Common.Name, 
                                   pattern = "American Shad") ~
                               "American Shad",
                        grepl(species, 
                                   pattern = "Herring|Alewife") ~
                               "River Herring",
                        grepl(species, 
                                   pattern = "American Shad") ~
                               "American Shad"),
         farthest = replace_na(farthest, 0) + 1,
         farthest2 = case_when(farthest %in% 2:4 ~ 2, 
                               farthest == 5 ~ 3,
                               TRUE ~ 1)) %>%
  mutate(tag_year = year(tag_time),
         far_tag = ifelse(farthest2 == 1, 1, 0),
         far_weir = ifelse(farthest2 == 2, 1, 0),
         far_pass = ifelse(farthest2 == 3, 1, 0)) %>%
  select(detect_year, tag_year, 
         tag_short, tag_time, sp,starts_with("n_"), 
         far_tag, far_weir, far_pass) %>%
  arrange(sp, far_tag, far_weir, far_pass)
                        

stairs %>%
  group_by(sp) %>%
  summarise(far_tag = mean(far_tag),
            far_weir = mean(far_weir),
            far_pass = mean(far_pass)) %>%
  mutate(tag_c = far_tag + far_weir + far_pass,
         weir_c = far_weir + far_pass,
         pass_c = far_pass) %>%
  select(sp, ends_with("_c")) %>%
  pivot_longer(cols = 2:4, values_to = "p") %>%
  mutate(stage = c(1,2,3,1,2,3),
         stage2 = rep(c("Tagged", 
                        "Reached\nladder", 
                        "Passed\nladder"), 2),
         n = p*c(51, 51, 51, 162, 162, 162),
         stage2 = forcats::fct_reorder(stage2, stage)) %>%
  ggplot() +
  geom_col(aes(x = stage2, y = n),
           position = position_dodge(width = 1)) +
  facet_wrap(~ sp, scales = "free") +
  theme_bw() +
  theme(text = element_text(size = 14),
        strip.background = element_rect(fill = "white")) +
  labs(x = "", y = "Number of individuals")

# ggsave("figures/stairstep.png", width = 7, height = 5, dpi = 400)

```

# Histogram of fish counts by hour
```{r}
# video time series
v_time_hist <- 
  read_xlsx("other/2 IFW Video Data Golden Master 11_2_21.xlsx",
               sheet = "Grouped") %>%
  select(Date, time = Time, species = Species) %>%
  mutate(year = year(Date),
         # note: time read in as decimal hours (a prop. of 24 hr)
         time = as.numeric(as.character(time))*24,
         hour = trunc(time),
         y2013 = case_when(year == 2013 ~ "2013",
                                  TRUE ~ "2014-2019")) %>%
  filter(species %in% c("American Shad", "River Herring spp.")) %>%
  arrange(species, time)


v_time_hist %>%
  group_by(species, y2013, hour) %>%
  summarize(n = length(hour),
            .groups = "drop") %>%
  group_by(species, y2013) %>%
  mutate(p = n/sum(n)) %>%
  ggplot() +
  geom_col(aes(x = hour, y=p),
                 binwidth = 1, 
                 color = "darkgray",
                 closed = "left",
                 breaks = c(0:24)) +
  facet_grid(y2013~species) +
  theme_bw() +
  theme(text = element_text(size = 14),
        strip.background = element_rect(fill = "white")) +
  labs(x = "Hour", y = "Proportion of individuals")

ggsave("figures/fish_counts_by_hour_histograms.png", 
       width = 7, height = 4, dpi = 400)
 

```

# % hours with pings during each day

```{r}

# Function to plot # pings/day by antenna to check antenna coverage
# make a function to plot pings/day by antenna for a given year
plot_antenna_pct_hrs_active <- function(ping_data){
  
  # get year
  the_year <- as.numeric(unique(ping_data$year))
  
  # make data frame of 1st tagging dates
    first_day <- tags %>% 
    mutate(year = year(tag_time),
           jul = yday(tag_time)) %>%
    group_by(year) %>%
    summarise(first_day = min(jul)) %>%
    mutate(date = c("Apr 27", "Apr 26", "Apr 14", "Apr 13", 
                    "Apr 6", "Apr 12", "Apr 10", "Apr 10"))
    
  plot_data <- ping_data %>%
    mutate(day = yday(datetime),
           hr = lubridate::hour(datetime)) %>%
    group_by(day, hr, antenna) %>%
    tally() %>%
    arrange(antenna, day, hr) %>%
    mutate(n1 = ifelse(n >= 3, 1, 0)) %>%
    group_by(antenna, day) %>%
    summarize(pct = 100*sum(n1)/24,
              .groups = "drop")

   antenna_plot <-
    ggplot(plot_data) +
    geom_col(aes(x = day, y = pct)) +
    geom_vline(xintercept = filter(first_day, 
                                   year == the_year)$first_day,
               color = "red", lty = 2) +
    facet_wrap(~antenna) +
    theme_bw() +
    scale_x_continuous(limits = c(91,182),
                       breaks = c(32, 60, 91, 121, 152, 
                                  182, 213, 244, 274),
                       labels = c("Feb", "", "Apr", "May", "Jun", 
                                  "Jul", "Aug", "", "Oct")) +
    theme(text = element_text(size = 15)) +
      labs(y = "% of hours active", x = "",
           title = as.character(the_year))
  
 ggsave(here("figures", paste0(stringr::str_sub(year(ping_data$date[1])), 
                                "_antenna_pct_hrs_active.png")), 
         height = 4, width = 7, dpi = 400)
  
  return(antenna_plot)
}

# Plot pings/day by antenna to check antenna coverage
plot_antenna_pct_hrs_active(data2012)
plot_antenna_pct_hrs_active(data2013)
plot_antenna_pct_hrs_active(data2014)
plot_antenna_pct_hrs_active(data2015)
plot_antenna_pct_hrs_active(data2016)
plot_antenna_pct_hrs_active(data2017)
plot_antenna_pct_hrs_active(data2018)
plot_antenna_pct_hrs_active(data2019)
```


# Fish passage time relative to sunrise/sunset
```{r}

read_nonleap <- function(year){
# year <- 2015

file_name <- paste0("sun", as.character(year), ".txt")

sun_file1 <- read.table(here("data",
                           "sun",
                           file_name), 
                      skip = 9,
                      nrows = 28
                      ) %>%
  select(1,8:15) %>%
  rename(day = 1,
         apr.sr = 2,
         apr.ss = 3,
         may.sr = 4,
         may.ss = 5,
         jun.sr = 6,
         jun.ss = 7,
         jul.sr = 8,
         jul.ss = 9
         )

sun_file2 <- read.table(here("data",
                           "sun",
                           file_name), 
                      skip = 37,
                      nrows = 2
                      ) %>%
  select(1,6:13) %>%
  rename(day = 1,
         apr.sr = 2,
         apr.ss = 3,
         may.sr = 4,
         may.ss = 5,
         jun.sr = 6,
         jun.ss = 7,
         jul.sr = 8,
         jul.ss = 9
         )

sun_file3 <- read.table(here("data",
                           "sun",
                           file_name), 
                      skip = 39,
                      nrows = 1
                      ) %>%
  select(1,6:9) %>%
  rename(day = 1,
         may.sr = 2,
         may.ss = 3,
         jul.sr = 4,
         jul.ss = 5
         )

sun_file <- sun_file1 %>%
  bind_rows(sun_file2) %>%
  bind_rows(sun_file3) %>%
  pivot_longer(cols = apr.sr:jul.ss, 
               values_to = "time") %>%
  mutate(month = substr(name, 1, 3),
         type = substr(name, 5, 7)) %>%
  pivot_wider(id_cols = c(month, day),
              values_from = time,
              names_from = type) %>%
  filter(is.na(sr)==F) %>%
  mutate(sr.h = as.numeric(substr(sr, 1,1)),
         sr.m = as.numeric(substr(sr, 2,3)),
         ss.h = as.numeric(substr(ss, 1,2)),
         ss.m = as.numeric(substr(ss, 3,4)),
         year = year,
         mo = case_when(month == "apr" ~ 4,
                        month == "may" ~ 5,
                        month == "jun" ~ 6,
                        month == "jul" ~ 7),
         datetime.sr = ymd_hm(paste0(year,"-",mo, "-", day, " ", 
                              sr.h,":",sr.m)),
         datetime.ss = ymd_hm(paste0(year,"-",mo, "-", day, " ", 
                              ss.h,":",ss.m)),
         ord = yday(datetime.sr)) %>%
  arrange(datetime.sr)

return(sun_file)

}

```

