---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---
# Packages
Load packages, functions, and tagging data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(wesanderson)
library(lubridate)
source("scripts/plot_by_tag_function.R")
source("scripts/plot_by_antenna_function.R")
source("scripts/fishsum_function.R")
`%notin%` <- Negate(`%in%`)

# read in tagging data
tags <- read.csv("other/tagging_data_11_2_21.csv") %>%
  filter(Tag.Number != "No Tag Number",
         Tag.Number != "-") %>%
  mutate(tag_short = str_sub(Tag.Number, - 9, - 1),
         tag_time = mdy_hm(Date.and.Time, tz = "America/New_York"))
```
# Compile receiver data
Data are compiled, filtered, and cleaned within these scripts. Anthony should review to ensure all filtering was OK.
```{r}
source("scripts/compile_pings_2012.R")
source("scripts/compile_pings_2013.R")
source("scripts/compile_pings_2014.R")
source("scripts/compile_pings_2015.R")
source("scripts/compile_pings_2016.R")
source("scripts/compile_pings_2017.R")
source("scripts/compile_pings_2018.R")
source("scripts/compile_pings_2019.R")
```
# Export cleaned, compiled data for inspection
```{r}
write.csv(data2012, 
          "output/2012_clean_detection_data.csv", row.names=F)
write.csv(data2013, 
          "output/2013_clean_detection_data.csv", row.names=F)
write.csv(data2014, 
          "output/2014_clean_detection_data.csv", row.names=F)
write.csv(data2015, 
          "output/2015_clean_detection_data.csv", row.names=F)
write.csv(data2016, 
          "output/2016_clean_detection_data.csv", row.names=F)
write.csv(data2017, 
          "output/2017_clean_detection_data.csv", row.names=F)
write.csv(data2018, 
          "output/2018_clean_detection_data.csv", row.names=F)
write.csv(data2019, 
          "output/2019_clean_detection_data.csv", row.names=F)
```
# Plot pings by tag to identify which belong to fish
```{r}
# plot pings by tag number to check and identify fish
plot_by_tag(data2012)
plot_by_tag(data2013)
plot_by_tag(data2014)
plot_by_tag(data2015)
plot_by_tag(data2016)
plot_by_tag(data2017)
plot_by_tag(data2018)
plot_by_tag(data2019)
```
# Plot pings by tag to identify which belong to fish
```{r}
plot_by_antenna(data2012)
plot_by_antenna(data2013)
plot_by_antenna(data2014)
plot_by_antenna(data2015)
plot_by_antenna(data2016)
plot_by_antenna(data2017)
plot_by_antenna(data2018)
plot_by_antenna(data2019)
```
# Extracting & summarizing fish data
extract and summarize data from fish tags
```{r}
fish2012_sum <- fishsum(data2012)
fish2013_sum <- fishsum(data2013)
fish2014_sum <- fishsum(data2014)
fish2015_sum <- fishsum(data2015)
fish2016_sum <- fishsum(data2016)
fish2017_sum <- fishsum(data2017)
fish2018_sum <- fishsum(data2018)
fish2019_sum <- fishsum(data2019)
fish_sum <- fish2019_sum %>%
  bind_rows(fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum) %>%
  arrange(year, Common.Name, tag_time)

# write fish summary csv files for review
# write.csv(fish_sum, "output/fish_sum.csv", row.names = F)

# remove unneeded objects
rm(fish2019_sum, fish2018_sum, fish2017_sum, fish2016_sum,
            fish2015_sum, fish2014_sum, fish2013_sum, 
            fish2012_sum)
```
# Format fish data for "passage proportion" analysis
```{r}
# subset target species and create 1/0 passage variable (did the fish reach A4)
f <- tags %>%
  select(tag_short, Common.Name, tag_time) %>%
  filter(Common.Name %in% c("American Shad",
                            "River Herring",
                            "Blueback Herring")) %>%
  left_join(select(fish_sum, year, tag_short, first_ping_A4), by = "tag_short") %>%
  filter(!duplicated(tag_short)) %>% # remove 3 duplicates (inter-year returns)
  mutate(passage = case_when(is.na(first_ping_A4) ~ 0,
                               TRUE ~ 1),
         year = case_when(is.na(year) ~ year(tag_time),
                          TRUE ~ year))

cbind.data.frame(
  trials = aggregate(f$passage, by = list(f$year), length),
  successes = aggregate(f$passage, by = list(f$year), sum)[,2],
  prop = aggregate(f$passage, by = list(f$year), mean)[,2]
  )

shad_data <- list(
  D = 1,
  N = nrow(filter(f, Common.Name == "American Shad")),
  L = length(unique(filter(f, Common.Name == "American Shad")$year)),
  y = filter(f, Common.Name == "American Shad")$passage,
  ll = as.numeric(as.factor(filter(f, Common.Name == "American Shad")$year)),
  x = rbind(rep(1, nrow(filter(f, Common.Name == "American Shad"))))
)


```
# Defining the Stan model
Modified from here:
https://mc-stan.org/docs/2_28/stan-users-guide/hierarchical-logistic-regression.html
```{r}
cat(file = "scripts/the_stan_model.stan", 
"
data {
  int<lower=1> D;
  int<lower=0> N;
  int<lower=1> L;
  array[N] int<lower=0, upper=1> y;
  array[N] int<lower=1, upper=L> ll;
  array[N] row_vector[D] x;
}
parameters {
  array[D] real mu;
  array[D] real<lower=0> sigma;
  array[L] vector[D] beta;
}
model {
  for (d in 1:D) {
    mu[d] ~ normal(0, 100);
    for (l in 1:L) {
      beta[l, d] ~ normal(mu[d], sigma[d]);
    }
  }
  for (n in 1:N) {
    y[n] ~ bernoulli(inv_logit(x[n] * beta[ll[n]]));
  }
}
"
)
```

```{r}

library(rstan)
rstan_options(auto_write = TRUE)

stan(file = "scripts/the_stan_model.stan",
     model_name = "shad passage",
     data = shad_data)
```

