---
title: "fish_pass"
author: "Mike Allen"
date: "12/7/2021"
output: html_document
---
# Packages
Load packages, functions, and tagging data
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(wesanderson)
library(lubridate)
source("scripts/plot_by_tag_function.R")
source("scripts/plot_by_antenna_function.R")
source("scripts/fishsum_function.R")
`%notin%` <- Negate(`%in%`)

# read in tagging data
tags <- read.csv("other/tagging_data_11_2_21.csv") %>%
  filter(Tag.Number != "No Tag Number",
         Tag.Number != "-") %>%
  mutate(tag_short = str_sub(Tag.Number, - 9, - 1),
         tag_time = mdy_hm(Date.and.Time, tz = "America/New_York"))
```
# Compile receiver data
Data are compiled, filtered, and cleaned within these scripts. Anthony should review to ensure all filtering was OK.
```{r}
source("scripts/compile_pings_2012.R")
source("scripts/compile_pings_2013.R")
source("scripts/compile_pings_2014.R")
source("scripts/compile_pings_2015.R")
source("scripts/compile_pings_2016.R")
source("scripts/compile_pings_2017.R")
source("scripts/compile_pings_2018.R")
source("scripts/compile_pings_2019.R")
```
# Export cleaned, compiled data for inspection
```{r}
write.csv(data2012, 
          "output/2012_clean_detection_data.csv", row.names=F)
write.csv(data2013, 
          "output/2013_clean_detection_data.csv", row.names=F)
write.csv(data2014, 
          "output/2014_clean_detection_data.csv", row.names=F)
write.csv(data2015, 
          "output/2015_clean_detection_data.csv", row.names=F)
write.csv(data2016, 
          "output/2016_clean_detection_data.csv", row.names=F)
write.csv(data2017, 
          "output/2017_clean_detection_data.csv", row.names=F)
write.csv(data2018, 
          "output/2018_clean_detection_data.csv", row.names=F)
write.csv(data2019, 
          "output/2019_clean_detection_data.csv", row.names=F)
```
# Plot pings by tag to identify which belong to fish
```{r}
# plot pings by tag number to check and identify fish
plot_by_tag(data2012)
plot_by_tag(data2013)
plot_by_tag(data2014)
plot_by_tag(data2015)
plot_by_tag(data2016)
plot_by_tag(data2017)
plot_by_tag(data2018)
plot_by_tag(data2019)
```
# Plot pings by tag to identify which belong to fish
```{r}
plot_by_antenna(data2012)
plot_by_antenna(data2013)
plot_by_antenna(data2014)
plot_by_antenna(data2015)
plot_by_antenna(data2016)
plot_by_antenna(data2017)
plot_by_antenna(data2018)
plot_by_antenna(data2019)
```
# Extracting & summarizing fish data
extract and summarize data from fish tags
```{r}
fish2012_sum <- fishsum(data2012)
fish2013_sum <- fishsum(data2013)
fish2014_sum <- fishsum(data2014)
fish2015_sum <- fishsum(data2015)
fish2016_sum <- fishsum(data2016)
fish2017_sum <- fishsum(data2017)
fish2018_sum <- fishsum(data2018)
fish2019_sum <- fishsum(data2019)
```
# write fish summary csv files for review
```{r}
write.csv(fish2012_sum, "output/fish2012_sum.csv", row.names = F)
write.csv(fish2013_sum, "output/fish2013_sum.csv", row.names = F)
write.csv(fish2014_sum, "output/fish2014_sum.csv", row.names = F)
write.csv(fish2015_sum, "output/fish2015_sum.csv", row.names = F)
write.csv(fish2016_sum, "output/fish2016_sum.csv", row.names = F)
write.csv(fish2017_sum, "output/fish2017_sum.csv", row.names = F)
write.csv(fish2018_sum, "output/fish2018_sum.csv", row.names = F)
write.csv(fish2019_sum, "output/fish2019_sum.csv", row.names = F)

```
# analysis
```{r}



```

